---
title: 'The DataViz battle: Plotly vs ggplot2'
description: What can you do with plotly vs. ggplot2 and how do they compare on a simple chart?
author: Bryan Whiting
date: 2017-02-10
output:
  distill::distill_article:
    self_contained: false
categories: 
  - dataviz
  - r
  - ggplot
  - plotly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F)
library(plotly)
library(ggplot2)
```

R users fall in love with ```ggplot2```, the growing standard for data visualization in R. The ability to quickly vizualize trends, and customize just about anything you'd want, make it a powerful tool. Yet this week, I made a discovery that may reduce how much I used ```ggplot2```. Enter ```plot_ly()```. 

For this post, I assume that you have a working knowledge of the ```dplyr``` (or ```magrittr```) and ```ggplot2``` packages. I caveat that this post is backed with only 4-5 hours using ```plotly()```, so some statements here may not be fully vetted.

Plotly and ```ggplot2``` are inherently for different purposes. ```plotly``` allows you to quickly create beautiful, reactive D3 plots that are particularly powerful in websites and dashboards. You can hover your mouse over the plots and see the data values, zoom in and out of specific regions, and capture stills. Here's a basic histogram:


```{r, message=F,warning= F}
set.seed(1)
library(plotly)
plot_ly(x = ~ rnorm(1000), type = "histogram")
```

After a brief dabble this week in ```plotly```, I realized quickly the many advantages that ```plotly``` has over ```ggplot2```. 

Several initial impressions:

* Plotly handles multiple wide data columns. I always find it annoying that to color different series in ggplot2, your data had to be in long format. Granted, it takes one simple ```melt()``` command to get the data into wide format.
* Plotly also handles long format (see below).
* Customizing the layout (plot borders, y axis) is easier.
* Customizing the legend is easier (in ```ggplot2``` I've wanted to remove just one series, which isn't always easy).
* Documentation is better in Plotly.
* Plotly syntax is very intuitive (learning how ```aes()``` in ```ggplot2``` works is tricky at first)
* Plotly also works for Python, Matlab, and Excel, among other languages.
* It's very easy to add new series and customize them (one line, one scatter, and one bar, for example)
* You can use other fonts (which is possible in ```ggplot2```, but I've never gotten to work on my Windows machine)
* You can toggle series on and off by clicking the series name in the legend

Benefits of ```ggplot2``` over ```plotly```:

* Facet wrapping is very easy in ```ggplot2```. (I think you have to do [subplots](https://plot.ly/r/subplots/) in ```plotly```.)
* ```ggplot2``` is probably quicker for exploratory analysis.


# Converting ggplot2 into plotly

An additional benefit of ```plotly``` is that you can convert your ```ggplot()``` graphs into a ```plotly``` object.

```{r,message=F,warning=F}
library(ggplot2)
p <- qplot(x = rnorm(1000), geom = "histogram")
p
```

Then, invoking the ```ggplotly(p)``` command, we see the transformation:

```{r,message=F,warning=F}
ggplotly(p)
```

A draw back of ```ggplotly()``` is that if you do refined customization (like putting your legend on the bottom of the graph), ```ggplotly()``` doesn't seem to pick this up by default. 

```{r,message=F,warning=F}
# ggplot with legend on the bottom
p <- qplot(
  data = iris,
  x = Sepal.Width,
  y = Sepal.Length,
  geom = "point",
  color = Species
) +
  theme(legend.position = "bottom")
p
# Plotly doesn't pick up the legend change
ggplotly(p)
```

But since Plotly also saves to an object, you can use the `%>%` notation to pipe and add additional plotting commands. This is similar to the `+` operator in
`ggplot()`.

```{r,message=F,warning=F}
p <- qplot(
  data = iris,
  x = Sepal.Width,
  y = Sepal.Length,
  geom = "point",
  color = Species
) +
  theme(legend.position = "bottom")
p2 <- ggplotly(p)
# Use the plotly layout() command for legend customization
p2 %>% layout(legend = list(orientation = "h"))
```

The legend doesn't do exactly what we want, but you can manipulate the legend location manually using x and y coordinates. The `orientation = 'h'` setting in the [docs](https://plot.ly/r/legend/) puts the legend on the bottom for default `plot_ly()` objects. Graphing the same series, we see the legend at the bottom:


```{r, message=F, warning=F}
plot_ly(iris,
  x = ~Sepal.Width,
  y = ~Sepal.Length,
  type = "scatter",
  mode = "markers",
  color = ~Species
) %>%
  layout(legend = list(orientation = "h"))
```

(You notice the Plotly X-axis title can get cut off[^2], so let's put that `+1` to `ggplot2`.)

[^2]: This may depend on your screen.

Plotly seems very intuitive relative to `ggplot2` in doing layout customization. Things that took me many iterations on StackOverflow to figure out, like adding a black line on y = 0, are built in to Plotly. 

```{r, message=F, warning=F}
p <- plot_ly(iris,
  x = ~Sepal.Width,
  y = ~Sepal.Length,
  type = "scatter",
  mode = "markers",
  color = ~Species
)
# Put legend on bottom, change the x-axis range, and turn on the x-axis line. 
# Also, make the zeroline visible, and turn it red.
p <- p %>% layout(
  legend = list(orientation = "h"),
  xaxis = list(
    zeroline = T, # Turns x = 0 on
    zerolinecolor = "red", # colors x = 0 red
    showline = T, # Shows xaxis border line
    range = c(-2, 7)
  )
)
# Or, save parameters into a list. Use new fonts (a huge plus)
f1 <- list(
  family = "Arial, sans-serif",
  size = 18,
  color = "lightgrey"
)
yax <- list(
  title = "Sepal length",
  titlefont = f1
)
p %>% layout(yaxis = yax)
```



Things I'd like to further explore:

- You can [export static plotly](https://plot.ly/r/static-image-export/) images out to file. My hypothesis is that Plotly images take longer to generate than `ggplot2`. So if I'm mass producing 30,000 plots (which I had to do last month), which is the faster approach? I would assume `ggplot2`.



# Plotly in RShiny Dashboards

The goal in learning Plotly was for me was to eliminate the Excel-VBA dashboard I created using for my manager. Excel has (some) benefits over `ggplot2` static charts: you can easily hover your mouse over a series to see the data value, and most industry users know how to manage an Excel axes. Grated, you can build in an RShiny widget to allow the user to control the axes, but Excel comes with that knowledge base built-in. `ggvis` allows for the powerful library of Google charts, but I think for a reactive dashboard, `plotly` is a great way to go[^1].

[^1]: Of course, this is relative to the chart you're trying to make.

So Plotly solved the Excel problem for me. Now my manager can click and zoom to the parts of the graph that are interesting, and hover the mouse to see the values. Just use `renderPlotly()` instead of `renderPlot()` in the `server.R` file, and `plotlyOutput()` instead of `plotOutput()` in the `ui.R` file.

More info here: [RShiny and Plotly](https://plot.ly/r/shiny-tutorial/) 

# RShiny vs Plotly Dashboards

Both RShiny and Plotly allow for creating dashboards. Plotly allows you to build [dashboards](https://plot.ly/r/dashboard/) as well. If you're just interested in only visualizing charts and trends, Plotly dashboards seem like the way to go. But to build reactivity into your dashboard (like subsetting your sample, changing date ranges, etc.), RShiny still seems like the more customizable solution. 

# Final thoughts

Overall, it seems that `ggplot2` is quicker to build and find what you want. 
With facet wrapping, the `qplot()` command, and `ggsave()`, you can whip
something up fast. Plotly is better for dashboards, as you can interact with 
the plots. I feel like Plotly has a better syntax and documentation, and so it 
may be easier to get a basic plot to look how you want it to. But `ggplot2` 
seems to have more advanced features, so if you want to get into refined 
customization, you may want to stick with `ggplot2`. They're both great, 
and serve different purposes, but I'll be using `plotly` for my RMarkdown 
and RShiny visualizations going forward.
