---
title: "Goals"
comments: false
listing:
  type: table
  contents: goals
  fields: [date, title, description]
  sort: 
    - "date desc"
    - "title asc"
  sort-ui: [date, title]
  filter-ui: [date, title]
  page-size: 30
  grid-item-border: true
  feed: false
  categories: cloud
execute:
  echo: false
  warning: false
  message: false 
---

::::{.callout-tip collapse="true"}
## Public accountability is the greatest motivator.

:::{.callout-tip collapse="true"}
## These goals are correlated with, but not causal to happiness. 

They're [anti-sad](https://journals.sagepub.com/doi/10.1177/1948550614568161?papetoc=), but pro-happy comes not from checking off boxes but by building [meaningful relationships](https://hbr.org/2010/07/how-will-you-measure-your-life) with God, self, family, and friends.

Why do I measure? Because whenever things aren't going well I often attribute it to a lacking of one of these things. But I don't keep this checkbox list to make me happy. Happiness is perspective and relationships. Purpose. So why this list? Well, they're the foundation of a guided, structured, and prosperous life.

- Studying scriptures and praying teaches me who I am and what my purpose is here on earth. It also gives me perfect role models to follow (Christ's example in the scriptures, and God's influence on me during prayer). What better way to guide your rudder (as [Christensen](https://hbr.org/2010/07/how-will-you-measure-your-life) says), than to be influenced by God on a daily basis. They say you're the average of the 5 people you spend the most time with. Why not make the number 1 person you spend your most valuable time to be God? (I don't have to spend 5 hours. 5 very meaningful minutes can correct my day.)

- Writing about my study is my little attempt to share God's light with the world. Imagine what social media would look like if everyone shared an insight from Marcus Aurelius, Buddha, Ghandi, Jesus, Mohammad, on a daily basis and how that insight affected their perspective? Be the change you want to see in the world, they say.

- Writing gratitude rewires my brain. It causes me to see things I wouldn't. Why not just say things I'm grateful for during prayer or a gratitude journal? I dunno. Those things don't help me for some reason. It always feels forced (I've tried it dozens of ways). Somehow, writing in public, gives me a great sense of "I mean it." Helps me think about that thing. Give it real thought.

- Physical health. Strength helps my bones. The effects of weight lifting are seen over decades, not months. I need to take better care of my bones. Same with cardio. They both energize today, but they prevent a whole host of bad things tomorrow and beyond. The effects are compounding - in both directions. Physical health also drives me to eat better. Eating better doesn't cause me to exercise, but exercising caueses me to eat better. Less snacks. 

- Meditating helps me train my mind. It's a moment where I think about something I want to improve on. Less reactivity as a dad to kids misbehaving. More optimism. More appreciation for the body I have, the health I enjoy, the freedom I have in this world.

:::

::: {.callout-tip collapse="true"}
## 2024 Goals: Read, hike, pray, meditate, workout, give thanks daily, study scripture daily

### Objective

- Deepen my relationship with and connection to God, self, family, and friends.
- Enjoy intellectual growth and historical understanding
- Have fun.

### KPIs 

Reading:

- 52 books. 12 fiction. 12 biographies. 12 business/self help. 4 religious. And 12 free choice (history, philosophy, career/data-ai). See [Reading List](lists/reading-list.md).

Daily Habits:

- Pray
- Meditate
- Workout (strength and cardio)
- Give thanks (write a post)
- Study scriptures (write a post)

Weekly Habits: Family Home Evening with my kids (Come Follow Me study + activity)

Monthly: Date night.

Quarterly: Attend the temple - a religious building (separate from church).

Fun:

- 52 hikes (1x/wk)
- Call a friend
- 4 weekend trips/vacations to new destinations and parks

:::
::::


```{r}
#| label: prep-data
library(dplyr)
library(ggplot2)
library(purrr)
library(tidyr)
library(gt)


# list_fields = c('categories', 'body-zones', 'exercies')

extract_yaml_to_df <- function(file) {
  # file = "christianity/three-little-pigs-christ.md"  
  # cat("Reading file", file)
  file_content <- readLines(file)
  yaml_lines <- grep("---", file_content)
  yaml_start <- yaml_lines[1] + 1
  yaml_end <- yaml_lines[2] - 1
  # yaml content
  yaml_list <- file_content[(yaml_start):(yaml_end)]
  yaml_text = paste(yaml_list, collapse = "\n")
  yaml_data <- yaml::read_yaml(text=yaml_text)
  yaml_data
  # Anything that's a vector should be listed and collapsed
  list_fields <- names(yaml_data)[sapply(yaml_data, function(x) is.vector(x) && length(x) > 1)]
  for(field in list_fields){
    yaml_data[field] = paste(unlist(yaml_data[field]), collapse=", ")
  }
  return(tibble::as_tibble(yaml_data))
}

# add gratitude
dirs <- c('goals', 'gratitude', "christianity")
df <- tibble::tibble()
for(dir in dirs){
  files <- list.files(path = dir, pattern = "*.md", full.names = TRUE)
  for(file in files){
    # print(file)
    # convert to df
    df_new <- extract_yaml_to_df(file) 
    df_new$folder = dir
    df <- bind_rows(df, df_new)
  }
}


df$date <- lubridate::parse_date_time(df$date, orders = c("ymd", "ymd HMS", "ymd HM", "ymdHMS", "ymdHM")) 
df$date <- as.Date(df$date)

# add missing columsn
for(col in c('run-dist', 'hike-dist', 'jumprope-time')){
  if(!(col %in% colnames(df))){
    df[[col]] <- 0
  }
}
# fill in missing defaults, characters are set to 'x' by default in _day journal
df <- df %>%
  mutate(across(where(is.character), ~ifelse(. == 'x', NA, .)))

# Create "cardio" and "strength" flags
df <- df %>% 
  mutate(
    post_cfm_study = stringr::str_detect(categories, 'come follow me'),
    post_gratitude = folder == 'gratitude',
    pelo = ifelse(`pelo-time` > 0, 1, 0),
    cardio = ifelse(`pelo-time` > 0 | `run-dist` > 0 | `hike-dist` > 0 | `jumprope-time` > 0, 1, 0),
    strength = ifelse(!is.na(exercises) >= 1, 1, 0)
  ) %>%
  filter(title != "test")

# df2 = daily df
df2 <- df %>%
  select(
    date, pray, meditate, cardio, strength, post_cfm_study, post_gratitude
  ) %>% 
  mutate(across(-date, as.numeric)) %>%
  mutate(across(-date, ~replace_na(., 0))) %>%
  group_by(date) %>%
  summarize_all(max) %>%
  ungroup() %>%
  arrange(desc(date))
```

::: {.panel-tabset}

##### Habits

```{r}
#| label: fig-goals
#| fig-height: 2
#| fig-width: 7
#| fig-align: center
#| fig-cap: "Goals"
df3 <- df2 %>% 
  rowwise() %>%
  mutate(total_goals = sum(pray, meditate, strength, cardio, post_cfm_study, post_gratitude)) %>%
  ungroup() %>%
  select(date, total_goals) 


df_dates <- bind_rows(data.frame(date = seq(as.Date("2024-01-01"), as.Date("2024-12-31"), by = "days"))) %>%
  mutate(day_of_month = lubridate::day(date)) %>%
  mutate(month_type = ifelse(lubridate::month(date) %% 2 == 0, 1, 0))

df4 <- left_join(df_dates, df3, by = 'date') %>%
  mutate(#week_of_year = lubridate::week(date),
         week_start_date = lubridate::floor_date(date, "week"),
         day_of_week = lubridate::wday(date, label = TRUE)) %>%
  mutate(total_goals = replace_na(total_goals, 0))



# p1 <- ggplot(df4, aes(x = week_start_date, y = day_of_week)) +
#   geom_tile(aes(fill = as.factor(month_type)), show.legend = FALSE) +
#   scale_fill_manual(values = c("0" = "white", "1" = "lightgray")) + 
#   theme_minimal()

# p2 <- ggplot(df4, aes(x = week_start_date, y = day_of_week)) +
#   geom_tile(aes(fill = total_goals), color = "white", width = 5, height = 0.9) +
#   scale_fill_gradient(low = "white", high = "darkgreen", limits = c(1, 6)) +
#   theme_minimal()

# library(gtable)
# g1 <- ggplotGrob(p1)
# g2 <- ggplotGrob(p2)

# # Make sure the sizes match
# g2$widths <- g1$widths
# g2$heights <- g1$heights

# # Overlay the plots
# g <- rbind(g1, g2, size = "first")
# g <- gtable::gtable_add_grob(g, g2$grobs, t = 1, l = 1, z = Inf)

# # Draw the combined plot
# grid.newpage()
# grid.draw(g)

# p1 + annotation_custom(ggplotGrob(p2), xmin = min(df4$week_start_date), xmax = max(df4$week_start_date), ymin = min(df4$day_of_week), ymax = max(df4$day_of_week))




ggplot(df4, aes(x = week_start_date, y = day_of_week)) +
  theme_minimal() +
  # background shading - nearly works.
  # geom_tile(aes(color=as.factor(month_type)), show.legend = FALSE) + 
  # geom_tile(aes(color=as.factor(month_type))) + 
  # scale_color_manual(values = c("white", "lightgray")) +
  # add titles for goals
  # geom_tile( aes(color = total_goals), color = "white", width = 5, height = 0.9) +
  geom_tile(data=df4, aes(fill = total_goals), color = "white", width = 5, height = 0.9) + 
  scale_fill_gradient(low = "white", high = "darkgreen", limits = c(0, 6)) +
  labs(x = "", y = "", fill='Habits\nAccomplished', title = "Daily Success") +
  scale_x_date(limits = as.Date(c("2023-12-25", "2024-12-31")), labels = scales::date_format("%b")) +
  scale_y_discrete(limits = rev(levels(df4$day_of_week))) +
  theme(axis.text.x = element_text(angle = 0, hjust = 0),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        axis.ticks = element_blank())
```

```{r}
#| label: fig-goals-table
#| fig-cap: Recent streak
#| fig-align: center
df2_recent <- df2 %>% 
  arrange(desc(date)) %>%
  head(14) %>%
  select(
    `Day` = date,
    `📖` = post_cfm_study,
    `🙏🏻` = post_gratitude,
    `🛐` = pray,
    `🚴🏻‍♂️` = cardio,
    `🏋🏼‍♂️` = strength,
    `🧘🏼‍♂️` = meditate,
  ) 

# Assuming your data frame is named 'data'
# First, convert the 'Day' column to the desired format
df2_recent$Day <- format(as.Date(df2_recent$Day), "%a %b %d")

df2_recent %>%
  gt(id='mygt', rowname_col = "Day") %>%
  data_color(
    method = "numeric",
    palette = c("white", "darkgreen"),
    domain = c(0, 1)
  ) %>%
  tab_header(
    title = "Daily Habits"
  ) |>
  tab_style(
    style = list(
      cell_text(color = "transparent")
    ),
    locations = cells_body(
      columns = c(2:7)  # replace with the actual column indices or names
    )
  ) %>%
  cols_align(align = "center") %>%
  tab_style(
    style = cell_borders(sides = "right", color = "white", weight = px(1)),
    locations = cells_body(columns = everything())
  ) %>%
  cols_width(
    Day ~ px(100),
    everything() ~ px(50),
  ) %>%
  tab_source_note(
    md("📖: Study, 🙏🏻: Gratitude, 🛐: Pray, 🚴🏻‍♂️: Cardio, 🏋🏼‍♂️: Strength, 🧘🏼‍♂️: Meditate")
  )
```

##### Records
```{r}
#| label: fig-peloton
#| fig-cap: Peloton Records
#| fig-align: center
df %>% 
  filter(pelo == 1) %>%
  filter(date >= Sys.Date() - 60) %>%
  group_by(`pelo-time`) %>%
  filter(`pelo-kj` == max(`pelo-kj`)) %>%
  select(`pelo-time`, `pelo-kj`, `pelo-avg`, date, `workout-notes`) %>%
  ungroup() %>%
  mutate(`90%` = paste0("", .9 * `pelo-kj`, "KJ\n(", .9 * `pelo-avg`, "avg)")) %>%
  mutate(`80%` = paste0("", .8 * `pelo-kj`, "KJ (", .8 * `pelo-avg`, "avg)")) %>%
  mutate(`70%` = paste0("", .7 * `pelo-kj`, "KJ (", .7 * `pelo-avg`, "avg)")) %>%
  gt() %>%
  cols_width(
    `90%` ~ px(70),
    `80%` ~ px(70),
    `70%` ~ px(70),
    # everything() ~ px(150),
  ) %>%
  cols_label(
    `pelo-time` = "Time",
    `pelo-kj` = "KJ",
    `pelo-avg` = "Avg W",
    `workout-notes` = "Notes",
    `date` = "Day"
  ) %>%
  tab_header(title = "Bike Records")

```

:::

***
